---
title: "Homework 4"
output: html_document
date: "2/13/2025"
---

Kaustubh Duddala
[Github](https://github.com/kaustubhduddala/StatisticsAndDataScience/tree/main/SDS%20315/Homework%203)

```{r, include=FALSE}
library(ggplot2)
library(tidyverse)
library(mosaic)
```

## Problem 1    

Are the observed data (70 flagged trades out of 2021) consistent with the SEC’s null hypothesis that, over
the long run, securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other
traders?

#### Iron Bank
```{r, message=FALSE, echo=FALSE}
# A
p1_iron_bank_fraud <- rbinom(n = 10000, size = 2021, prob = 0.024)
p1_pvalue = mean(p1_iron_bank_fraud >= 70)

cat("p value:", p1_pvalue)

hist(p1_iron_bank_fraud, main="Distribution of simulated flagged trades under null hypothesis", xlab = "Number of flagged trades")

```

[write up]

## Problem 2

Are the observed data for Gourmet Bites consistent with the Health Department’s null hypothesis
that, on average, restaurants in the city are cited for health code violations at the same 3% baseline rate?

#### Health Inspections
```{r, message=FALSE, echo=FALSE}
# A
p2_health <- rbinom(n = 100000, size = 50, prob = 0.03)
p2_pvalue = mean(p2_health >= 8)

cat("P value:", p1_pvalue)

hist(p2_health, main="Distribution of simulated health code violations under null hypothesis", xlab = "Number of health code violations")

```

[write up]

## Problem 2

Are the observed data for Gourmet Bites consistent with the Health Department’s null hypothesis
that, on average, restaurants in the city are cited for health code violations at the same 3% baseline rate?

#### Health Inspections
```{r, message=FALSE, echo=FALSE}
# A
p2_health <- rbinom(n = 100000, size = 50, prob = 0.03)
p2_pvalue = mean(p2_health >= 8)

cat("P value:", p1_pvalue)

hist(p2_health, main="Distribution of simulated health code violations under null hypothesis", xlab = "Number of health code violations")

```

[write up]