---
title: "Homework 8"
output:
  html_document: default
  pdf_document: default
date: "4/6/2025"
---

Kaustubh Duddala

[Github](https://github.com/kaustubhduddala/StatisticsAndDataScience/tree/main/SDS%20315/Homework%208)

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 999)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(mosaic)
library(MatchIt)
```

```{r, include=FALSE}
data = read_csv("creatinine.csv")
```

## Question 1



#### A

```{r, message=FALSE, echo=FALSE}

crea_lm <- lm(creatclear ~ age, data = data)
beta_age <- coef(crea_lm)["age"]
beta_0 <- coef(crea_lm)["(Intercept)"]
predicted_clearance_55 <- beta_0 + beta_age * 55

```

The estimated creatinine clearance for a 55-year-old individual is about `r round(predicted_clearance_55, digits = 2)` ml/minute. This was calculated by building a linear regression model using the creatinine.csv data, extracting the intercept and the coefficient for age, and then applying the formula: `creatclear = intercept + (age coefficient × 55)`. This allows us to estimate the expected clearance rate for any given age.

#### B

The estimated value of B₁ (the age coefficient or slope) from our linear model is about `r round(beta_age, 2)`. This means that for each additional year of age, the expected creatinine clearance decreases by about `r abs(round(beta_age, 2))` mL/minute.

#### C

```{r, message=FALSE, echo=FALSE}

residual_40 <- 135 - (beta_0 + beta_age * 40)
residual_60 <- 112 - (beta_0 + beta_age * 60)

```
For the 40-year-old, the observed creatinine clearance is about `r round(residual_40,2)` mL/minute higher than what the model predicts. For the 60-year-old, the observed clearance is `r round(residual_60,2)` mL/minute above the expected value. This suggests that the 40-year-old (with a clearance rate of 135) is relatively healthier for their age compared to the 60-year-old (with a rate of 112). These residuals were calculated by subtracting the model’s predicted value from the observed value for each individual:
for the 40-year-old, `135 - (intercept + age coefficient × 40) = 12`.



## Question 2
```{r, message=FALSE, echo=FALSE}

covid_data <- read_csv("covid.csv")

italy_data <- covid_data %>% filter(country == "Italy")
spain_data <- covid_data %>% filter(country == "Spain")

fit_italy <- lm(log(deaths) ~ days_since_first_death, data = italy_data)
rate_italy <- coef(fit_italy)["days_since_first_death"]
italy_dbl_time <- log(2) / rate_italy

set.seed(123)
italy_boot <- do(10000) * lm(log(deaths) ~ days_since_first_death, data = resample(italy_data))
italy_rate_ci <- confint(italy_boot)
italy_dbl_ci <- confint(log(2) / italy_boot$days_since_first_death)

fit_spain <- lm(log(deaths) ~ days_since_first_death, data = spain_data)
rate_spain <- coef(fit_spain)["days_since_first_death"]
spain_dbl_time <- log(2) / rate_spain

set.seed(123)
spain_boot <- do(10000) * lm(log(deaths) ~ days_since_first_death, data = resample(spain_data))
spain_rate_ci <- confint(spain_boot)
spain_dbl_ci <- confint(log(2) / spain_boot$days_since_first_death)

ggplot(covid_data, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line() +
  labs(
    title = "COVID-19 Daily Deaths by Country",
    x = "Days Since First Death",
    y = "Number of Deaths",
    color = "Country"
  ) 


```

For both Italy and Spain, we modeled daily COVID-19 deaths using an exponential growth model:  
`log(deaths) = intercept + growth_rate × days_since_first_death`  
The growth rate is the slope, and doubling time is log(2) divided by the growth rate, and 10,000 bootstrap samples are used to estimate 95% confidence intervals.

Italy: 
  - Growth rate: `r round(rate_italy, 2)`
  - Doubling time: `r round(italy_dbl_time, 2)` days
  - 95% CI for growth rate: (`r round(quantile(italy_boot$days_since_first_death, 0.025), 2)`, `r round(quantile(italy_boot$days_since_first_death, 0.975), 2)`)
  - 95% CI for doubling time: (`r round(quantile(log(2)/italy_boot$days_since_first_death, 0.025), 2)`, `r round(quantile(log(2)/italy_boot$days_since_first_death, 0.975), 2)`)
  
Spain:
  - Growth rate: `r round(rate_spain, 2)`
  - Doubling time: `r round(spain_dbl_time, 2)` days
  - 95% CI for growth rate: (`r round(quantile(spain_boot$days_since_first_death, 0.025), 2)`, `r round(quantile(spain_boot$days_since_first_death, 0.975), 2)`)
  - 95% CI for doubling time: (`r round(quantile(log(2)/spain_boot$days_since_first_death, 0.025), 2)`, `r round(quantile(log(2)/spain_boot$days_since_first_death, 0.975), 2)`)
  
Spain’s deaths grew faster and doubled more quickly than Italy’s during this period.



## Question 3

```{r, message=FALSE, echo=FALSE}

milk_data <- read_csv("milk.csv")

milk_model <- lm(log(sales) ~ log(price), data = milk_data)

elasticity_estimate <- coef(milk_model)["log(price)"]

set.seed(123)
n_boot <- 10000
boot_elasticities <- replicate(n_boot, {
  sample_indices <- sample(nrow(milk_data), replace = TRUE)
  boot_sample <- milk_data[sample_indices, ]
  coef(lm(log(sales) ~ log(price), data = boot_sample))["log(price)"]
})

elasticity_ci <- quantile(boot_elasticities, c(0.025, 0.975))
```


the estimated price elasticity of demand for milk was found by fitting a linear regression to the log-transformed sales and price data, using the model:
`log(sales) = intercept + elasticity × log(price)`
The elasticity coefficient (the slope) quantifies the percentage change in sales for a 1% change in price. To assess the uncertainty of the estimate, 10,000 bootstrap samples were used to calculate a 95% confidence interval.

Estimated elasticity: `r round(elasticity_estimate, 2)`

95% Confidence Interval: (`r round(elasticity_ci[1], 2)`, `r round(elasticity_ci[2], 2)`)

This means that a 1% increase in the price of milk is associated with an estimated `r abs(round(elasticity_estimate, 2))`% decrease in the quantity of milk sold.