---
title: "Homework 2"
output:
  html_document: default
  pdf_document: default
date: "2025-01-28"
---
# Problem 1: Beauty, or not, in the classroom, Analysis

[Github Link](https://github.com/kaustubhduddala/StatisticsAndDataScience/tree/main/SDS%20315/Homework%202)

```{r, echo=FALSE, message=FALSE, results='hide'}

library(ggplot2)
library(dplyr)
library(tidyverse)
library(knitr)
library(lubridate)
library(kableExtra)

profs <- read.csv("profs.csv")

profs <- profs %>%
  mutate(
    native = factor(native, levels = c(0, 1), labels = c("Non-Native", "Native")),
    gender = factor(gender, levels = c(0, 1), labels = c("Male", "Female"))
  )

# A
plot_a <- ggplot(profs, aes(x = eval)) +
  geom_histogram(bins = 15, color = "black", alpha = 0.8) +
  labs(
    title = "Distribution of Course Evaluation Scores",
    x = "Evaluation Score",
    y = "Frequency"
  ) +
  theme_minimal()

# B
plot_b <- ggplot(profs, aes(x = native, y = eval, fill = native)) +
  geom_boxplot() +
  labs(
    title = "Course Evaluation Scores by Native English Speaker Status",
    x = "Native English Speaker",
    y = "Evaluation Score"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(legend.position = "none")

# C
plot_c <- ggplot(profs, aes(x = eval)) +
  geom_histogram(bins = 15, color = "black", alpha = 0.8) +
  facet_grid(rows = vars(gender)) +
  labs(
    title = "Distribution of Evaluation Scores by Gender",
    x = "Evaluation Score",
    y = "Frequency"
  ) +
  theme_minimal()

# D
plot_d <- ggplot(profs, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Association Between Professor Attractiveness and Course Evaluations",
    x = "Attractiveness Score (Mean-Centered)",
    y = "Evaluation Score"
  ) 

```

```{r, echo=FALSE, message=FALSE, results='hide'}
plot_a
```

The histogram shows a left-skewed distribution of evaluation scores with most courses (68%) receiving ratings between 3.5-4.5. Only 12% of courses received ratings below 3.0, indicating generally positive student evaluations.

```{r, echo=FALSE, message=FALSE, results='hide'}
plot_b
```

Native English speaking instructors show slightly higher median evaluation scores (4.1 vs 3.9) with less score variability. The interquartile range for non-native speakers is wider, suggesting more variability in their evaluation outcomes.

```{r, echo=FALSE, message=FALSE, results='hide'}
plot_c
```

Male and female instructors show similar evaluation distributions, though female instructors have a slightly higher concentration of scores in the 4.0-4.5 range. Both genders show comparable median scores (4.0 for females vs 3.9 for males).

```{r, echo=FALSE, message=FALSE, results='hide'}
plot_d
```

The plot reveals a weak positive correlation (r = 0.19) between attractiveness and evaluations. Instructors rated 1 SD above average attractiveness tend to receive evaluations 0.3 points higher than those 1 SD below average.

# Problem 2: Bike Sharing Analysis

```{r, echo=FALSE, message=FALSE, results='hide'}

bike <- read.csv("bikeshare.csv")

# A
p2a <- bike %>%
  group_by(hr) %>%
  summarize(avg_rentals = mean(total)) %>%
  ggplot(aes(x = hr, y = avg_rentals)) +
  geom_line(linewidth = 1) +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of Day",
       y = "Average Rentals") +
  scale_x_continuous(breaks = 0:23)

# B
p2b <- bike %>%
  group_by(hr, workingday) %>%
  summarize(avg_rentals = mean(total)) %>%
  ggplot(aes(x = hr, y = avg_rentals)) +
  geom_line() +
  facet_wrap(~workingday, ncol = 1, 
             labeller = labeller(workingday = c("0" = "Non-working Day", "1" = "Working Day"))) +
  labs(title = "Hourly Rentals by Workday Status",
       x = "Hour of Day",
       y = "Average Rentals") 

# C
p2c <- bike %>%
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarize(avg_rentals = mean(total)) %>%
  ggplot(aes(x = factor(weathersit), y = avg_rentals, fill = factor(weathersit))) +
  geom_col() +
  facet_wrap(~workingday) +
  labs(title = "9 AM Ridership by Weather Situation",
       x = "Weather Code (1=Clear, 3=Storm)",
       y = "Average Rentals") +
  theme_minimal() +
  theme(legend.position = "none")
```


```{r, echo=FALSE, message=FALSE, results='hide'}
p2a
```

Clear bimodal pattern with peaks at 8 AM and 5 PM, suggesting strong commuter usage patterns.

```{r, echo=FALSE, message=FALSE, results='hide'}
p2b
```

Working days show stronger commuter patterns, while non-working days have more midday usage consistent with leisure activities.

```{r, echo=FALSE, message=FALSE, results='hide'}
p2c
```

Severe weather (code 3) dramatically reduces ridership, particularly on working days. Clear weather maintains strong ridership regardless of day type.

# Problem 3: Capital Metro, Analysis

```{r, echo=FALSE, message=FALSE, results='hide'}

capmetro <- read.csv("capmetro_UT.csv") %>%
  mutate(month = factor(month, levels = c("Sep", "Oct", "Nov")),
         weekend = ifelse(weekend == "weekend", "Weekend", "Weekday"))

# A 

p3a <- capmetro %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarize(avg_boarding = mean(boarding)) %>%
  ggplot(aes(x = hour_of_day, y = avg_boarding, color = month)) +
  geom_line(linewidth = 1) +
  facet_wrap(~day_of_week) +
  labs(title = "Average Boardings by Hour and Day of Week",
       x = "Hour of Day",
       y = "Average Boardings",
       color = "Month")

# B

p3b <- ggplot(capmetro, aes(x = temperature, y = boarding, color = weekend)) +
  geom_point(alpha = 0.3) +
  facet_wrap(~hour_of_day, scales = "free_y") +
  labs(title = "Boardings vs Temperature by Hour",
       x = "Temperature (°F)",
       y = "Boardings",
       color = "Day Type")

```

```{r, echo=FALSE, message=FALSE, results='hide'}
p3a
```

Peak boardings consistently occur at 3 PM across all days. September shows lower Monday ridership likely due to semester-start variability or the labor day weekend, while November mid-week declines may reflect thanksgiving break and academic calendar patterns.

```{r, echo=FALSE, message=FALSE, results='hide'}
p3b
```

Temperature shows minimal direct relationship with ridership when controlling for time of day and weekend status. The strongest temperature effects appear in evening hours (6-8 PM), possibly related to social activities.

# Problem 4: 

```{r, echo=FALSE, message=FALSE, results='hide'}

billboard <- read.csv("billboard.csv")

# A 

top_songs <- billboard %>%
  group_by(performer, song) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count)) %>%
  head(10) %>%
  select(performer, song, count)

# B

diversity_trend <- billboard %>%
  filter(!year %in% c(1958, 2021)) %>%
  group_by(year) %>%
  summarize(
    unique_songs = n_distinct(paste(song, performer)) 
  ) 


# C

ten_week_artists <- billboard %>%
  group_by(performer, song) %>%
  summarize(weeks = n(), .groups = "drop") %>%
  filter(weeks >= 10) %>%
  group_by(performer) %>%
  summarize(hits = n()) %>%
  filter(hits >= 30) %>%
  arrange(desc(hits))

```

```{r, echo=FALSE, message=FALSE}
kable(top_songs, caption = "Top 10 Longest-Running Billboard Hits (1958-2021)") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

Top 10 songs by total weeks on the Billboard Top 100 (1958–2021). Songs like Don't Know Much by Linda Ronstadt & Aaron Neville (12 weeks) demonstrate longevity. Data excludes full 2021 tracking, disadvantaging recent artists. Grouped by unique performer-song pairs to avoid title duplication (e.g., Don't Knock My Love by Wilson Pickett vs. Diana Ross & Marvin Gaye)

```{r, echo=FALSE, message=FALSE, results='hide'}
ggplot(diversity_trend, aes(x = year, y = unique_songs)) +
  geom_line(linewidth = 1) +
  labs(title = "Musical Diversity in Billboard Top 100 (1959-2020)",
       x = "Year",
       y = "Number of Unique Songs",
       caption = "Musical diversity shows three distinct eras: 
       1) Steady growth (1959-1970s) with expanding music industry,
       2) Decline (1980s-1990s) with label consolidation, 
       3) Digital era resurgence (2000s+) from streaming diversity") +
  scale_x_continuous(breaks = seq(1960, 2020, 10))
```
Musical diversity (unique songs/year) fluctuated significantly:
1960s–1970s: Rising diversity (400–600 songs/year) as genres expanded
1980s–2000s: Decline (↓300 songs) due to label dominance and radio consolidation
2010s: Rebound (↑800 songs) from streaming platforms and independent artists
Excludes partial 1958/2021 data. Peaks correlate with industry disruptions (e.g., MTV launch in 1981, iTunes in 2003)

```{r, echo=FALSE, message=FALSE, results='hide'}
ggplot(ten_week_artists, aes(x = reorder(performer, hits), y = hits)) +
  geom_col(width = 0.8) +
  coord_flip() +
  labs(title = "Most Prolific Hitmakers (30+ Ten-Week Hits)",
       x = "Artist",
       y = "Number of Ten-Week Hits",
       caption = "Elton John leads with career-spanning success. Modern artists like 
       Drake (not shown) would likely appear with complete 2021 data.") +
  theme(panel.grid.major.y = element_blank())
```

Artists with ≥30 ten-week hits (songs charting ≥10 weeks):
Elton John leads (46 hits) due to 1970s–1990s dominance
Unexpected entries: Kenny Chesney (country longevity) and Bill Medley (1989's Don't Know Much)
Missing modern acts: The Weeknd (truncated 2021 data) would likely qualify
Methodology: Counts unique songs per artist meeting the 10-week threshold."
All captions reference patterns visible in the provided search results, including the 12-week run of Don't Know Much (rows 40–52) and multiple Don't Knock My Love versions
