---
title: "SDS 313 Week 2 - More Visualization Options"
output: html_document
date: "2024-09-05"
---

## More Visualization Options

Today we'll look at:

-   Adding layers to a graph
-   Exporting your graphs
-   The ggplot2 graphics package

But first let's start with the practice we missed last class, using your survey data from last week:

```{r}
#Import Survey Data:
survey <- read.csv('SDS313_StudentSurvey.csv')

```

### Group Practice

1.  Pick a variable that your group thinks might differ between those of you who are from Texas versus from outside of Texas. Describe and display the bivariate relationship.
2.  Post your code, graph, and descriptive statistics on your group's slide [here](https://docs.google.com/presentation/d/1dD7Mn2YNGz2INQRh5toxPwvBKScJUBfoEtmtloeLzKY/edit?usp=sharing).

```{r}
#Group practice plot:
fromTexas <- survey$Grow_up == "In Texas"
notFromTexas <- survey$Grow_up == "Outside of Texas"
texas <- survey$Grow_up
likesTacos <- survey$Tacos

barplot(table(survey$texas))


```

### Import Birthweight Dataset

We'll spend today's class using a real dataset that contains the birth weights (in grams) of a sample of babies born at a hospital in Springfield, MA.

```{r}
birthwt <- read.csv('birthwt.csv')
```

What code can we use to get familiar with the variables in this dataset?

```{r}
#Write some code to visualize the variables in this dataset:
smoke <- birthwt

```

## Layering

We can add additional layers to a graph to include more information about the data. Like using pch creatively. First run just the base plot:

```{r}
#Build base scatterplot
plot(birthwt$lwt,birthwt$bwt,main='Mothers Weight and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20,col='grey')
```

Now run the entire chunk below and compare to above to see what *points()* is doing:

```{r}
#Build base scatterplot
plot(birthwt$lwt,birthwt$bwt,main='Mothers Weight and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20,col='grey')

#Add another layer of points
points(birthwt$lwt,birthwt$bwt)
```

We can also layer on text boxes:

```{r}
#Build base scatterplot
plot(birthwt$lwt,birthwt$bwt,main='Mothers Weight and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20)

#Add text box at given coordinates
text(230,2000,'Write text here')
```

Or add lines:

```{r}
#Build base scatterplot
plot(birthwt$lwt,birthwt$bwt,main='Mothers Weight and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20)

#Add lines
abline(h=mean(birthwt$bwt),col='blue',lty=2)
abline(v=mean(birthwt$lwt),col='dark green',lwd=4)
abline(lm(birthwt$bwt~birthwt$lwt),col='red')
```

Or color different groups with subsetting. Let's practice subsetting again:

```{r}
#Replace the ????? to create smoking yes/no subsets
SmokeYes <- birthwt[birthwt$smoke == 'yes',]
SmokeNo <- birthwt[birthwt$smoke == 'no',]
```

Now we can color the subsets separately:

```{r}
#Scatterplot of one subset
plot(SmokeYes$lwt,SmokeYes$bwt,main='Mothers Weight and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20,col='orange')

#Add points for other subset
points(SmokeNo$lwt,SmokeNo$bwt,pch=20,col='blue')
```

Don't forget to add a legend:

```{r}
#Scatterplot of one subset
plot(SmokeYes$lwt,SmokeYes$bwt,main='Mothers Weight and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20,col='orange')

#Add points for other subset
points(SmokeNo$lwt,SmokeNo$bwt,pch=20,col='blue')

#Add legend
legend('bottomright',c('Smokers','Non-Smokers'),pch=c(20,20),col=c('orange','blue'), inset=.03, cex=.8)
```

## Exporting a graph

You can use RStudio's user-friendly buttons to export plots, or use code. If using code, exporting a graph takes 3 steps:

1.  Naming a file to write to.
2.  Running the graph function(s).
3.  Closing the file.

```{r}
#Where is this plot saved to...?
pdf('mygraph.pdf') 
plot(birthwt$lwt, birthwt$bwt) 
dev.off()
```

### Group Practice

1.  Is mother's age related to birth weight? Make a nice graph to display this relationship.
2.  Within your graph, come up with a way to highlight or in someway distinguish babies with low birth weight (defined as less than 2,500 grams).
3.  Export your graph using code and paste it on your group's slide [here](https://docs.google.com/presentation/d/1dD7Mn2YNGz2INQRh5toxPwvBKScJUBfoEtmtloeLzKY/edit?usp=sharing).

```{r}
#Group Practice:
age = birthwt$age
bwt = birthwt$bwt
lbwt = birthwt[birthwt$bwt < 2500,]
plot(age,bwt,main='Mothers Age and Birthweight',xlab='Mothers Pre-Pregnancy Weight (lbs)',ylab='Birthweight (g)',pch=20,col='orange')

points(lbwt$age, lbwt$bwt,pch=20, col="blue")

```

## ggplot Graphics

A very popular graphics package that can make it easier to create nice looking graphs is [ggplot2](https://ggplot2.tidyverse.org/index.html).

Make sure the package is already installed and then call it:

```{r}
library(ggplot2)
```

The code for ggplot functions looks a little different, but you'll see similarities with base R plots. Note however that the default graphs come out much nicer.

We start by defining a *ggplot()* object, which must take a data frame:

```{r}
#Define ggplot object with a dataframe
bw_ggplot <- ggplot(birthwt)
```

Then we "add" elements to it based on what type of graph we want:

```{r}
#Bar graph
bw_ggplot + geom_bar(aes(x=smoke))
```

```{r}
#Histogram
bw_ggplot + geom_histogram(aes(x=bwt))
```

```{r}
#Scatterplot
bw_ggplot + geom_point(aes(x=lwt,y=bwt))
```

We can change properties of the graph by **adding** more layers:

```{r}
#Add titles
bw_ggplot + geom_point(aes(x=lwt,y=bwt)) + labs(title='Mothers Weight and Birth Weight', x='Mothers Weight (lbs)',y='Birth Weight (g)')

#Change shape
bw_ggplot + geom_point(aes(x=lwt,y=bwt), pch=24)

#Adjust axis limits
bw_ggplot + geom_point(aes(x=lwt,y=bwt)) + xlim(c(50,300)) + ylim(c(0,6000))
```

If you want to get rid of classic "gray box" background:

```{r}
bw_ggplot + geom_point(aes(x=lwt,y=bwt)) + theme_classic()
```

One of the best things about ggplot is that it makes coloring or filling by any other variable very easy - ***no subsetting needed!***

```{r}
#Side-by-side bar graph
bw_ggplot + geom_bar(aes(x=ftv,fill=low), position = "dodge")
```

```{r}
#Grouped histogram
bw_ggplot + geom_histogram(aes(x=bwt,fill=smoke), alpha=.5, binwidth=250, position='identity')
```

```{r}
#Scatterplot with grouped points
bw_ggplot + geom_point(aes(x=lwt,y=bwt,col=smoke))
```

It plays well with **color brewer**:

```{r}
#Open library and view color palettes 
library(RColorBrewer)
display.brewer.all()
```

```{r}
#Side-by-side bar graph
bw_ggplot + geom_bar(aes(x=ftv,fill=low), position = "dodge") + scale_fill_brewer(type='seq',palette=1)

#Scatterplot with grouped points
bw_ggplot + geom_point(aes(x=lwt,y=bwt,col=smoke)) + scale_colour_brewer(type='qual',palette=2)
```

**Facet wraps** make splitting your graph by a categorical variable extremely easy:

```{r}
bw_ggplot + geom_histogram(aes(x=bwt),binwidth=250,col='black',fill='dark red')  + labs(title='Birthweights by Race',x='Birthweight (g)',y='Frequency') + theme(legend.position='bottom') + scale_fill_brewer(type='qual',palette=2) + facet_wrap(~race)
```

It can also produce **gradient coloring**:

```{r}
bw_ggplot + geom_point(aes(x=lwt,y=bwt,color=age)) + labs(title='Birthweight by Age and Mothers Weight',x='Mothers Weight (lbs)',y='Birthweight (g)')
```

Some additional references for ggplot:

-   [Function reference](https://ggplot2.tidyverse.org/reference/index.html)
-   [R Cookbook](http://www.cookbook-r.com/Graphs/)
-   [ggplot Cheat sheet](https://github.com/rstudio/cheatsheets/blob/main/data-visualization.pdf)

### Group Practice

Using the *StudentSurvey.csv* file from last week's class and the ggplot2 functions, complete the following in your groups:

1.  Make a histogram of the number of books our class has read in the past year. Include appropriate labels and a title. Try changing colors or the bin size.

2.  Make overlapping histograms of books read split by the video games variable. Try moving the legend around, or changing the color palette.

3.  Change the histograms into density curves.

4.  Pick three variables and make a graph(s) that investigates their relationship with how many books a student has read in the past year. Post your graph and code on your [group's slide](https://docs.google.com/presentation/d/1dD7Mn2YNGz2INQRh5toxPwvBKScJUBfoEtmtloeLzKY/edit?usp=sharing).

```{r}
survey <- read.csv('SDS313_StudentSurvey.csv')

```
