---
title: "Project 2: Analysis of the IMDb Top 1000 Worst Rated Titles "
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
date: "2024-11-14"
---

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
library(rvest)
library(stringr)
library(dplyr)
library(knitr)
```

# Introduction

This project will analyze the IMDb Top 1000 Worst Rated Titles dataset, which lists the lowest-rated movies and TV shows on IMDb with over 10,000 votes. This dataset is sourced from Kaggle, and it is updated daily by OctopusTeam. You can find the dataset [here](https://www.kaggle.com/datasets/octopusteam/imdb-top-1000-worst-rated-titles?resource=download).

I chose this dataset because it offers insights into the factors contributing to poor audience reception. By exploring patterns in genre, release year, and rating, we can investigate trends and identify potential reasons why certain titles fail to resonate with audiences. This analysis may also reveal insights about the correlation between genres and their ratings, shedding light on audience preferences.

![Image Source: Wikimedia Commons](https://upload.wikimedia.org/wikipedia/commons/6/69/IMDB_Logo_2016.svg)

## Dataset Import and Processing

The dataset’s usability rating on Kaggle (10/10) suggests it is ready for analysis with minimal processing required. It will be imported into R directly, and if necessary, minor reshaping will be done to facilitate analysis.

## Variables of Interest

Key variables I will investigate include:

**Genres**: To see if certain genres are overrepresented in low ratings.

**Average Rating**: To analyze the distribution of ratings.

**Number of Votes**: To explore if vote count correlates with ratings.

**Release Date**: To explore if certain release dates ended up more often on this list than others

This exploration will help identify potential trends and audience behaviors that influence IMDb ratings.

# Data Preparation

```{r, echo=FALSE, include=FALSE}

data <- as_tibble(read.csv("data.csv"))

```

The data was obtained from Kaggle, which provided an already well-structured and cleaned dataset and it was largely pre-processed, meaning there were no significant issues with missing values (NA), factors, or date formatting. I did a quick check to ensure the data types matched my analysis needs, but overall, the dataset was good to go as is.

One of the benefits of using a dataset from Kaggle is that it often comes with clear documentation and formatting, minimizing the usual challenges of data preparation, like handling inconsistencies or dealing with null values. This allowed me to jump directly into the analysis without needing to spend any additional time in processing.

# Variable Descriptions

```{r, echo=FALSE}

data <- data %>%
  select(title, genres, averageRating, numVotes, releaseYear)

genre_subset_data <- data %>%
  select(genres, averageRating, numVotes, releaseYear) %>%
  separate_rows(genres, sep = ", ")

variable_info <- tibble(
  Variable = c("genres", "averageRating", "numVotes", "releaseYear"),
  Type = c("Character", "Numeric", "Integer", "Integer"),
  Description = c(
    "The genre of the movie",
    "Average rating of the movie from user votes",
    "Total number of votes received for the movie",
    "Year the movie was released"
  )
)

kable(variable_info, caption = "Description of Variables in the Cleaned Dataset")

```

# Univariate Analyses

```{r, echo=FALSE}

ggplot(genre_subset_data, aes(x = genres)) +
  geom_bar(width = 1, alpha = 0.2, color = "purple", fill = "purple") +
  labs(title = "Frequency of Genres", x = "Genres", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, echo=FALSE}
ggplot(data, aes(x = averageRating)) +
   geom_histogram(binwidth = 0.1, alpha = 0.2, color = "purple") +
   labs(title = "Distribution of ratings", x = "Ratings", y = "Frequency")
```

```{r, echo=FALSE}
ggplot(data, aes(x = numVotes)) +
   geom_histogram(binwidth = 1000, alpha = 0.2, color = "purple") +
   labs(title = "Distribution of number of votes", x = "Number of Votes", y = "Frequency")

```

```{r, echo=FALSE}
ggplot(data, aes(x = releaseYear)) +
   geom_histogram(binwidth = 1, alpha = 0.2, color = "purple") +
   labs(title = "Distribution of release year", x = "Release year", y = "Frequency")

```

# Multivariate Analyses

```{r, echo=FALSE}

ggplot(data, aes(x = averageRating, y = numVotes)) +
  geom_point(alpha = 0.1, color = 'purple') +
  labs(title = "Average Rating vs Number of Votes",
       x = "Average Rating",
       y = "Number of Votes")
```

```{r, echo=FALSE}
ggplot(data, aes(x = releaseYear, y = averageRating)) +
  geom_point(alpha = 0.2, color = 'purple') +
  labs(title = "Release Year vs Average Rating",
       x = "Release Year",
       y = "Average Rating")
```

```{r, echo=FALSE}
ggplot(genre_subset_data, aes(x = genres, y = numVotes)) +
  geom_jitter(width = 0.3, alpha = 0.05, color = "purple") +
  labs(title = "Number of Votes by Genre",
       x = "Genre",
       y = "Number of Votes") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, echo=FALSE}
ggplot(genre_subset_data, aes(x = genres, y = averageRating)) +
  geom_jitter(width = 0.3, alpha = 0.2, color = "purple") +
  labs(title = "Average Rating by Genre",
       x = "Genre",
       y = "Average Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Reproducibility Resources

# Choice Elements

# Conclusion

# References
