---
title: "SDS 313 Week 2 - Visualizing and Describing Data"
output: html_document
date: "2024-09-03"
---

## Visualizing and Describing Data

In the next two classes, we'll be:

-   Plotting different types of data
-   Running univariate and bivariate descriptive statistics
-   Looking at the ggplot graphics package

We'll be working with *R Markdown* files during lectures throughout the rest of the semester (we'll learn to make our own in a couple of weeks).

Each file will contain text in addition to code "chunks," which act the same way as submitting code to the console.

Hit the run button (little green triangle) in the upper right to see this in action:

```{r}
#Run this chunk
5+2
```

### Import Dataset

The default working directory of any *R Markdown* file will be the folder where it is saved. So as long as your dataset is in the same folder, you don't need need to worry about file paths:

```{r}
med <- read.csv('medicaldata.csv')
```

What code can we use to get familiar with the variables in this dataset?

```{r}
#Write some code to visualize the variables in this dataset:

```

## Summarizing a Categorical Variable

A categorical variable defines membership in a group.

-   Describe with: *frequencies or proportions*
-   Display with: *bar graph*

```{r}
#Frequency table
table(med$Diabetic)

#Proportions
prop.table(table(med$Diabetic))
```

```{r}
#Bar graph
barplot(table(med$Diabetic))
```

If we want to label the categories:

```{r}
#Update dummy values with labels
med$Diabetic <- factor(med$Diabetic, labels=c('No','Yes'))

#See the difference?
table(med$Diabetic)
barplot(table(med$Diabetic))
```

We should also add some options:

```{r}
#Nicer looking bar graph
barplot(table(med$Diabetic),main='Frequency of Diabetic Status',xlab='Diabetic Status',ylab='Frequency',ylim=c(0,100),col=c('aquamarine','purple'))
```

Here are some other resources that can help make your base R plots look nicer:

-   [Graph parameters](http://www.statmethods.net/advgraphs/parameters.html)
-   [Color list](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)
-   [Nice color palettes](http://colorbrewer2.org/)

Update the bar graph with your two favorite colors. Try calling them with their hexidecimal code:

```{r}
#Bar graph with your colors:

```

## Summarizing a Numeric Variable

A numeric variable is a quantitative measurement.

-   Describe with:
    -   Center: *mean or median*
    -   Spread: *standard deviation or inner-quartile range (IQR)*
-   Display with: *histogram or boxplot*

```{r}
#Mean (average value)
mean(med$Cholesterol)

#Median (50th percentile)
median(med$Cholesterol)
```

```{r}
#Standard deviation (average distance to the mean)
sd(med$Cholesterol)

#Five-number summary (min, Q1, median, Q3, max)
fivenum(med$Cholesterol)

#Inner-quartile range (Q3-Q1)
IQR(med$Cholesterol)
```

When do we use which measures of center/spread? It depends on the shape of the distribution:

-   **Symmetric** distributions: mean, standard deviation
-   **Skewed** distributions: median, IQR

We'll need to visualize the distribution to determine its shape:

```{r}
#Basic histogram
hist(med$Cholesterol)

#Nicer looking histogram
hist(med$Cholesterol,main='Histogram of Cholesterol',xlab='Cholesterol (mg/dL)',col='dark green',xlim=c(0,2000),ylim=c(0,120))
```

```{r}
#Basic boxplot
boxplot(med$Cholesterol)

#Nicer looking boxplot
boxplot(med$Cholesterol,main='Boxplot of Cholesterol',ylab='Cholesterol (mg/dL)', pch=20)

#Horizontal boxplot
boxplot(med$Cholesterol,main='Boxplot of Cholesterol',xlab='Cholesterol (mg/dL)', pch=20, horizontal=TRUE)
```

### Group Practice

1.  Pick a variable from this dataset we haven't see yet (not diabetic status or cholesterol). Describe and display its distribution.
2.  Change the default color of your plot and update the title and axis labels.
3.  Go to your group's slide [here](https://docs.google.com/presentation/d/1dD7Mn2YNGz2INQRh5toxPwvBKScJUBfoEtmtloeLzKY/edit?usp=sharing) and write everyone's names at the top. Then copy/paste your graph onto your slide.

```{r}
#Group practice plot:

```

## Summarizing a Bivariate Relationship

### 2 Numeric Variables

-   Describe with: *correlation*
    -   *-1 \< r \< 1*
-   Display with: *scatterplot*

```{r}
#Pearson correlation coefficient
cor(med$BMI,med$BP)
```

```{r}
#Basic scatterplot
plot(med$BMI,med$BP)

#Nicer looking scatterplot
plot(med$BMI,med$BP,main='BMI and Blood Pressure',xlab='BMI',ylab='Blood Pressure (mmHg)',pch=20)
```

### 2 Categorical Variables

-   Describe with: *row or column proportions*
-   Display with: *two-way frequency table or grouped bar graph*

Let's label the edema status variable like we did diabetic:

```{r}
#Add labels to Edema variable
med$Edema <- factor(med$Edema, labels=c('No','Yes'))

#Two-way frequency table
table(med$Diabetic,med$Edema)
```

```{r}
#Proportion options for two-way table
prop.table(table(med$Diabetic,med$Edema))
prop.table(table(med$Diabetic,med$Edema),1)
prop.table(table(med$Diabetic,med$Edema),2)
```

```{r}
#Side-by-side grouped bar graph
barplot(table(med$Diabetic,med$Edema),main='Diabetes and Edema Status',ylab='Frequency',col=c('#018571','#b2abd2'),beside=T, xlab='Edema Status')
```

We can also "stack" are bar chart to look at proportions. What should we insert into the barplot() function instead of our table()?

```{r}
#Make "stacked" bar chart: replace ???? with the appropriate function
barplot(????(table(med$Diabetic,med$Edema),2),main='Diabetes and Edema Status', ylab='Proportion', col=c('#018571', '#b2abd2'), xlab='Edema Status')
```

### 1 Numeric and 1 Categorical Variable

-   Describe with: *compare center/spread of each group*
-   Display with: *grouped histogram or grouped boxplot*

```{r}
#Descriptive statistics by group
aggregate(BP~Diabetic, data=med, mean)
```

```{r}
#What about the standard deviations of each group?

```

Sub-setting practice:

```{r}
#Create subsets for each Diabetic group:
Diabetic <- 
NonDiabetic <- 
```

```{r}
hist(Diabetic$BP,main='Histogram of Blood Pressure',xlab='Blood Pressure (mmHg))',col='#31a35475',ylim=c(0,25),xlim=c(40,100))

hist(NonDiabetic$BP,col='#756bb175',add=T)
```

```{r}
boxplot(med$BP~med$Diabetic, main='Diastolic Blood Pressure',ylab='Blood Pressure (mg/dL)',xlab='Diabetic Status',names=c('No','Yes'),col=c('#0571b0','#ca0020'))
```

### Group Practice

1.  Pick a variable that might be related to Saliva Glucose Level (Glucose). Describe and display the bivariate relationship.
2.  Post your graph on your group's slide [here](https://docs.google.com/presentation/d/1dD7Mn2YNGz2INQRh5toxPwvBKScJUBfoEtmtloeLzKY/edit?usp=sharing).

```{r}
#Group practice plot:

```

*Before next class, make sure you have the following packages ready:*

```{r}
#Do you have these packages installed?
library(ggplot2)
library(RColorBrewer)
```
