---
title: "SDS 313 Week 7 - Reshaping Data"
output: html_document
date: "2024-10-08"
---

## Reshaping Data

Today, we'll:

-   Learn what characteristics make a dataset "tidy."
-   Reshape datasets wide-to-long and long-to-wide.

## The "Tidy" Data Format

We usually want the [following characteristics](https://d33wubrfki0l68.cloudfront.net/6f1ddb544fc5c69a2478e444ab8112fb0eea23f8/91adc/images/tidy-1.png) to be true for any dataset we are going to analyze:

-   Each variable is in its own column.
-   Each observation is in its own row.
-   Each value is in its own cell.

Such a format is referred to as *tidy*, as in the *tidyverse*.

Below is a set of example datasets, all storing the same information: the number of Tuberculosis cases documented by the WHO across three different countries in 1999 and 2000.

***Which of these tables is "tidy"?***

```{r}
#Call tidyverse library:
library(tidyverse)

#View the same data organized in different tables:
table1
table2
table3
table4a
table4b
```

## Reshaping "Untidy" Data

### Convert wide data to long with pivot_longer

We can take "wide" data and combine values of a single variable that are spread across multiple columns with *pivot_longer()*.

For example, let's take the "cases" that are spread out across two variables in *table4a* and gather them into the same column (thus making it longer).

```{r}
#Wide data:
table4a

#Provide existing variable names and define new variables:
table4a_long <- pivot_longer(table4a, cols = c('1999','2000'),  names_to = 'year', values_to = 'cases')
table4a_long
```

### Group Practice 1

1.  Reshape the information in table4b from from wide to long and call your new tibble "table4b_long".
2.  Merge the information in table4a_long and table4b_long into a single tibble.
3.  Check that your resulting tibble matches table1.

```{r}
#Group practice:
table4b

table4b_long <- pivot_longer(table4b, cols = c("1999","2000"), names_to = "year", values_to="cases")
table4b_long

table4ab_long <- full_join(table4b_long, table4a_long)
table4ab_long
```

### Convert long data to wide with pivot_wider

We can also reshape in the opposite direction, by spreading out long data to extract multiple variables that are defined in one column with values in another.

For example, let's take the cases and population "counts" in *table2* and put them in their own columns (thus making it wider):

```{r}
#Long data:
table2

#Provide existing variable names and define new variables:
table2_wide <- pivot_wider(table2, id_cols = c('country','year'), names_from = 'type', values_from = 'count')
table2_wide
```

### Separate

We can separate() different values existing in the same column:

```{r}
#Two values contained in one column:
table3

#Separate them into their own variables:
table3_sep <- separate(table3, rate, into=c("cases", "population"), sep="/")
table3_sep
```

Notice the types of our new variables are the same as the original (character). But an easy option let's us convert the new variables to numeric:

```{r}
#Allow new variable to default to number:
table3_sep <- separate(table3, rate, into=c("cases", "population"), sep="/", convert = TRUE)
table3_sep
```

You can even separate the values from a single variable at a given position within a string:

```{r}
#Pull out year from century:
table3

table3_year <- separate(table3, year, into = c("century", "year"), sep = 2, convert = TRUE)
table3_year
```

### Group Practice 2

The *Swifties_Songs.csv* dataset is contains various characteristics of Taylor Swift songs. Import it and complete the following:

1.  How many different albums are represented in this dataset?
2.  Reshape your subset so that each song appears in its own unique row. How many different songs are contained this dataset?
3.  Pick your two favorite albums and create a subset that contains only songs from those two albums.
4.  Pick an "attribute" (your choice) to investigate. Make a bivariate graph that compares that attribute between your two albums, use *group_by()* to calculate the relevant bivariate statistics, and paste your results on your [group slide](https://docs.google.com/presentation/d/1dD7Mn2YNGz2INQRh5toxPwvBKScJUBfoEtmtloeLzKY/edit?usp=sharing).

```{r}
#Import dataset:
setwd("~/Documents/UT/SDS 313/12 Reshaping, Reordering")
swifties <- read_csv('Swifties_Songs.csv')

# 1
nrow(table(swifties$album))

# 2
swiftiesWide <- pivot_wider(swifties, id_cols = c("name", "album"), names_from = "Attribute", values_from = "Value")
swiftiesWide

#3
swiftiesFavs <- filter(swifties, album == "1989" | album == "Red")
swiftiesFavs
swiftiesFavesWide <- pivot_wider(swiftiesFavs, id_cols = c("name", "album"), names_from = "Attribute", values_from = "Value")
swiftiesFavesWideGrouped <- group_by(swiftiesFavesWide, danceability)
swiftiesFavesWideGrouped
```
